In-silico molecular enrichment and clearance of the human intracranial space

### 2025-01-26 paper/supplementary_main.tex

We solve the models presented in~\Cref{sec:app:more_maths} numerically
using finite element methods. The following subsections provide more
details, model-by-model. \rami{Add a note on the convection dominated problem and the need for DG}

@@

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time and
a discontinuous Galerkin (DG) finite element method with upwinding in
space to accurately capture sharp boundary layers. \mk{Comment on Peclet number}

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
polynomials of total degree $\leqslant 1$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
porosity-weighted diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
\begin{equation} \label{eq:3d_dg_disc}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

@@

We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ is the unit outward normal vector to each facet. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
\begin{equation}
\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
\end{equation}
Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
tangential component is enforced weakly via interior penalization. For
convenience, we collect all facets exterior to the CSF space but not
on the (Neumann) boundary $\Gamma_{\rm AM-U}$ in the set
$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L} \cup \mathcal{F}_{\rm SSAS}$, where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
interface $\Gamma_{\rm LV}$ by $\mathcal{F}_{\rm LV}$,
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$. Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
  - \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
  \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\}\cdot \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where on exterior facets the average and jump operators take the one-sided values. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that 
\begin{subequations}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
\end{align}
\end{subequations}

@@

  watson1983diffusion, asgari2016glymphatic, keith2019dispersion,

@@

estimates~\cite{watson1983diffusion, keith2019dispersion}. More

@@

  al.~\cite{keith2019dispersion} to estimate local enhancement factors

@@

%% Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}. These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, keith2019dispersion}. Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.

@@

%% Further assuming unsteady dispersion, we follow Sharp et al \cite{keith2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient 
%% $$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
%% Finally, we account for the non-local nature of the dispersion mechanism and smooth the resulting field by solving a Helmholtz problem with $dP+i^2/\alpha_i^3$ as the right-hand side
%% \begin{equation}
%% - 10^{-4} \Delta R_i + R_i = dP_i^2/\alpha_i^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF}
%% \end{equation}
%% to obtain the locally averaged, spatially varying diffusion enhancement factor $R_i$. The diffusion enhancement factor $R$ in $D^{\mathrm{eff}} = D^{\mathrm{Gad}}( 1 +R)$ is then estimated by $R \approx R_{\rm{cardiac}} + R_{\rm{resp}}$.

@@

We assess the numerical accuracy and convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions. Specifically, we generate a sequence of three meshes (\Cref{fig:mesh_refinement}) with an increasing number of computational mesh vertices and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for the baseline model on each mesh, and compare the results across meshes with respect to a set of key quantities of interest. The mesh refinement employed here is adaptive in the sense that the coarser mesh is refined near expected sharp concentration gradients. This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.  
Similarly, we investigate the effect of the time step size by solving the intracranial transport model with different time steps: 1, 2, and 4 minutes.

Considering the mean tracer concentrations in the CSF, parenchyma and
arterial and venous PVS domain over the first 24\,h after injection,
we observe negligible changes with both mesh and time refinement
(\Cref{fig:mesh_convergence_concentrations} and
\Cref{fig:time_convergence_concentrations}). As an additional
verification step, we compute the mean and maximum dispersion
enhancement factor, the maximum CSF pressure and velocity in both the
cardiac-driven and CSF production-induced flow fields, and mean
concentrations at 3, 6, 12 and 24\,h for all mesh resolutions and time
steps (\Cref{fig:mesh_time_convergence}). While the maximum dispersion factor
increases by about 60\,\% from the low resolution to the standard mesh, it
stabilizes with the next refinement step. All other quantities change
with less than 10\,\% with mesh refinement, and less than 1\,\% with
time refinement. We thus conclude that the standard resolution mesh
and a time step of 2 min offer sufficient accuracy for our
simulations.

@@

>>>>

We solve the models presented in~\Cref{sec:app:more_maths} numerically
using finite element methods. The following subsections provide more
details, model-by-model.

@@

We consider the system of coupled 3D-1D solute transport equations
given by~\eqref{eq:3d_pde} and~\eqref{eq:1d_pde} with the interface
conditions~\eqref{eq:3d_interface}, the bifurcation
conditions~\eqref{eq:conc_continuity}
and~\eqref{eq:flux_conservation}, and the boundary
conditions~\eqref{eq:3d_bcs} and~\eqref{eq:1d_bcs}. We discretize
these equations using an implicit finite difference scheme in time and
a discontinuous Galerkin (DG) finite element method with upwinding in space to accurately capture sharp boundary layers. We remark that the transport in the CSF spaces is highly convection dominated, with an average Péclet number of 402 and maximum of 9542 (assuming a characteristic length of $10\,$cm and taking the increased diffusivity due to dispersion into account).

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
polynomials of total degree $\leqslant 1$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
porosity-weighted diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter (we set $\eta=1000$). To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E c\bm{u} \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
\begin{equation} \label{eq:3d_dg_disc}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

@@

We consider an H(div)-based finite element approximation of the incompressible Stokes equations~\eqref{eq:stokes} defined over $\Omega_{\rm CSF}$ with the boundary conditions~\eqref{eq:stokes_bcs}. Following\cite{hong2016robust}, we approximate the velocity field $\bm u$ and the pressure field $p$ with the following finite element spaces:
\begin{align*}
  \bm V_{h,g} &= \{ \bm v  \in H(\mathrm{div}, \Omega_{\mathrm{CSF}}): \;
  \bm v \vert_{E} \in \mathrm{BDM}^2(E), \; E \in \mathcal{T}_{\rm CSF}; \;
  \bm v \cdot \bm n = 0
  \text{ on } \Gamma_{\rm{pia}} \cup \Gamma_{\rm AM-L} \cup \Gamma_{\rm{SSAS}}, \;\; \bm v \cdot \bm n =  g %\frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}} 
  \text{ on } \Gamma_{\mathrm{LV}} \}  \\ 
  Q_h  &= \{q \in L^2(\Omega_{\rm CSF}): \; q \vert_E \in \mathrm{P}^1(E),
  \; E \in \mathcal{T}_{\rm CSF} \}. 
\end{align*}
Here, $H(\mathrm{div}, \Omega_{\rm{CSF}})$ is the space of $L^2(\Omega_{\rm{CSF}})$ vector fields with $L^2(\Omega_{\rm{CSF}})$ divergence, $\mathrm{BDM}^2$ is the Brezzi-Douglas--Marini element~\cite{brezzi1987mixed} of degree 2, and $\bm n$ is the unit outward normal vector to each facet. Given any vector $\bm v$, the normal and tangential components on each facet are denoted and given by 
\begin{equation}
\bm v_n = (\bm v \cdot \bm n) \bm n, \quad \bm v_t = \bm v - \bm v_n. 
\end{equation}
Since $\bm V_{h,g} \subset H(\mathrm{div}, \Omega_{\rm CSF})$, then $[\bm
  v_n] = 0 $ on $\mathcal{F}_{i, \rm CSF}$. Continuity in the
tangential component is enforced weakly via interior penalization. For
convenience, we collect all facets exterior to the CSF space but not
on the (Neumann) boundary $\Gamma_{\rm AM-U}$ in the set
$\mathcal{F}_{e} = \mathcal{F}_{\rm pia} \cup \mathcal{F}_{\rm LV}
\cup \mathcal{F}_{\rm AM-L} \cup \mathcal{F}_{\rm SSAS}$, where facets
lying on the pial interface $\Gamma_{\rm pia}$ are denoted by
$\mathcal{F}_{\rm pia}$, on the lower and upper outer (arachnoid) boundary
$\Gamma_{\rm AM-L}$ and $\Gamma_{\rm AM-U}$ by $\mathcal{F}_{\rm AM-L}$ and $\mathcal{F}_{\rm AM-U}$, on the inner ventricular
%on the boundary toward the spinal cord $\Gamma_{\rm SC}$ by $\mathcal{F}_{\rm SC}$
and on the boundary towards the spinal SAS $\Gamma_{\rm SSAS}$ by
$\mathcal{F}_{\rm SSAS}$. Now, define the form 
\begin{multline}
  \mathcal{A}_h(\bm u, \bm v)
  = \sum_{E \in \mathcal{T}_{\rm CSF}} \int_{E} \mu \nabla \bm u : \nabla \bm v
  - \sum_{F \in \mathcal{F}_{i, \rm CSF} \cup \mathcal{F}_{e}} \left (
  \int_{F} \mu \{\nabla \bm u\} \cdot \bm n_F \cdot [\bm v_t] 
  - %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
  \int_{F} \mu \{\nabla \bm v\}\cdot \bm n_F \cdot [\bm u_t]
  + %\sum_{F \in \mathcal{F}_{h,c} \cup \mathcal{F}_{\Gamma}}
\int_{F}   \frac{\sigma \mu}{h_F} [\bm u_t ] \cdot [\bm v_t] \right ),   
\end{multline}
where on exterior facets the average and jump operators take the one-sided values. We set the penalty parameter for the tangential continuity to be $\sigma = 20$. Our finite element discretization of the incompressible Stokes equations is then to find $(\bm u_h, p_h) \in \bm V_{h,g} \times Q_h$ with $g  = \frac{1}{|\Gamma_{\mathrm{LV}}|} u_{\mathrm{in}}$  such that 
\begin{subequations}
\begin{align}
  \mathcal{A}_h(\bm u_h, \bm v_h)
  + \sum_{F \in \mathcal{F}_{\rm AM-U}} \int_{F} R_0 (\bm u_h \cdot \bm n) (\bm v_h \cdot \bm n)
  - \int_{\Omega_{\rm CSF}} \nabla \cdot \bm v_h \, p_h  &= 0 \qquad \foralls \bm v_h \in \bm V_{h,0} \\ 
\int_{\Omega_{\rm CSF}} \nabla \cdot \bm u_h \, q_h  &= 0 \qquad \foralls q_h \in Q_h.
\end{align}
\end{subequations}

@@

  watson1983diffusion, asgari2016glymphatic, sharp2019dispersion,

@@

estimates~\cite{watson1983diffusion, sharp2019dispersion}. More

@@

  al.~\cite{sharp2019dispersion} to estimate local enhancement factors

@@

%% Taking the large spatial differences of pulsatile CSF flow into account, we estimate a spatially varying diffusion enhancement factor $R$ in the CSF-filled space $\Omega_{\rm CSF}$ by combining a computational model of cardiac- and respiration-induced pulsatile CSF flow with theoretical estimates for shear-augmented (Taylor) dispersion \cite{taylor1953dispersion, watson1983diffusion}. These estimates are are derived under simplifying assumptions on the problem geometry, the flow field and concentration profiles \cite{watson1983diffusion, sharp2019dispersion}. Therefore such an estimate is heuristic; however, our computational results qualitatively agree with observed data.

@@

%% Further assuming unsteady dispersion, we follow Sharp et al \cite{sharp2019dispersion} and estimate each enhancement factor $R_{\rm cardiac}$ and $R_{\rm resp}$ from the non-dimensionalized pressure gradient 
%% $$R_{i} \approx dP_{i}^2 / \alpha_{i}^3, \quad  \text{where} \quad  dP=\frac{|\nabla p^{\rm up}_{i}|}{\omega_{i} \mu / h} \quad \text{for} \quad i = \{\mathrm{cardiac}, \mathrm{resp}\}.$$
%% Finally, we account for the non-local nature of the dispersion mechanism and smooth the resulting field by solving a Helmholtz problem with $dP+i^2/\alpha_i^3$ as the right-hand side
%% \begin{equation}
%% - 10^{-4} \Delta R_i + R_i = dP_i^2/\alpha_i^3 \quad \text{ on } \Omega_{\rm CSF} \qquad \text{ and } \quad \nabla R_i \cdot \bm n = 0, \quad \text{ on } \partial \Omega_{\rm CSF}
%% \end{equation}
%% to obtain the locally averaged, spatially varying diffusion enhancement factor $R_i$. The diffusion enhancement factor $R$ in $D^{\mathrm{eff}} = D^{\mathrm{Gad}}( 1 +R)$ is then estimated by $R \approx R_{\rm{cardiac}} + R_{\rm{resp}}$.

@@

We assess the numerical accuracy and convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions. Specifically, we generate a sequence of three meshes (\Cref{fig:mesh_refinement}) with an increasing number of computational mesh vertices and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for the baseline model on each mesh, and compare the results across meshes with respect to a set of key quantities of interest. The mesh refinement employed here is adaptive in the sense that the coarser mesh is manually refined near expected sharp concentration gradients. This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.  
Similarly, we investigate the effect of the time step size by solving the intracranial transport model on the standard resolution mesh with different time steps: 1, 2, and 4 minutes.

Considering the mean tracer concentrations in the CSF, parenchyma and
arterial and venous PVS domain over the first 24\,h after injection,
we observe negligible changes with both mesh and time refinement
(\Cref{fig:mesh_convergence_concentrations} and
\Cref{fig:time_convergence_concentrations}). As an additional
verification step, we compute the mean and maximum dispersion
enhancement factor, the maximum CSF pressure and velocity in both the
cardiac-driven and CSF production-induced flow fields, and mean
concentrations at 3, 6, 12 and 24\,h for all mesh resolutions and time
steps (\Cref{fig:mesh_time_convergence}). While the maximum dispersion factor
increases by about 60\,\% from the low resolution to the standard mesh, it
stabilizes with the next refinement step. All other quantities change
with less than 10\,\% with mesh refinement, and less than 1\,\% with
time refinement. We thus conclude that the standard resolution mesh and a time step of 2 min offer sufficient accuracy for our simulations, and remark that all reported results are obtained with the standard resolution mesh.

@@

\paragraph{Predictions on overall tracer spreading patterns}

\begin{table}
\centering
\begin{tabular}{c|c|c|c|c}
    \toprule
   quantity of interest  & Watts et al\cite{watts2019measuring}  & Ringstad et al\cite{ringstad2018brain} & Eide and Ringstand \cite{eide2024functional} & ours (vasomotion model)\\
   \midrule
    Spinal transit time &  1 - 3 hours & 20 min &  $13\pm6.3\,$ min & N/A \\
    Time of peak amount (CSF) & 10 - 15 hours & 6–9 hours & & 2 hours \\
    Time of peak amount (parenchyma) & 10 - 26 hours & 24 hours & N/A& 12 - 24 hours \\
    Peak amount (CSF) &  N/A & N/A & N/A & 0.5 mmol \\
    Peak amount (parenchyma) & N/A & N/A& N/A & 0.09 mmol \\
    injected tracer quantity &  1.0 mmol & 0.5 mmol & 0.5 mmol & 0.5 mmol (cranially) \\
    Peak concentration (CSF) & 0.5 mmol/l &  N/A&  N/A & 1.4 mmol/l\\
    Peak concentration (parenchyma) & 0.1 mmol/l & N/A & N/A & 0.4 mmol/l \\
    FTA in MCA1 & N/A & N/A &  37.8 ± 47.0 min &  1:48 / 1:24 hours \\
    FTA in MCA2 & N/A & N/A & 53.1 ± 50.5 min & 2:12 / 1:48 hours \\
    FTA in ACA2 & N/A & N/A & 48.3 ± 46.1 min & 3:00 / 3:12 hours \\
    \bottomrule
\end{tabular}
\caption{Comparison of literature values and model predictions on overall tracer spreading patters and timing}
\label{tab:spreading comparison}
\end{table}
